---
title: "Useful R packages"
author: "Jake Butts"
date: "3/11/2022"
output: html_document
---

```{r message=F}
library(tidyverse)

theme_set(theme_minimal())

set.seed(1)

knitr::opts_chunk$set(message=F)
```

```{r}
library(janitor)

colnames(iris)

#clean_names parses letter cases and separators to a consistent format (snake_format is default, but can be modified)

colnames(janitor::clean_names(iris))
```

```{r}
library(stringr)

#lets create some number strings
strings <- c(" 219 733 8965",
             "329-293-8753 ",
             "banana",
             "595 794 7569",
             "387 287 6718", "233.398.9187  ",
             "482 952 3315",
             "239 923 8115 and 842 566 4692",
             "Work: 579-499-7527",
             "$1000",
             "Home: 543.355.3679")

#we can then define pattern(s) for what a phone number looks like: 

phone_pattern <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})"

#str_extract will clean and pull values that fit this pattern, and return other values as NA. 
str_extract(strings, phone_pattern)
```

```{r}
#lubridate is incredibly useful for parsing and extracting information from date information. 

library(lubridate)

my_date = ymd(20200529) #ymd is just one of the patterns lubridate will accept. 
my_date_2= mdy(05292020)

my_date
my_date_2

#whatever format you specifiy, lubirdate will parse and return in Year-Month-Day. 

month(my_date) #we can pull specific components of our date

leap_year(my_date) #we can find out if its a leap year or not.
```

```{r}
#forcats is a great package for those of us who often work with catagorical data/variables. 

library(forcats)

fav_colors = sample(colors(), 25) #sample 25 colors from r's color pallete
counts = round(runif(25, 5, 100)) #get 25 rounded numbers from a uniform distribution between 5 and 100, call these counts

cols = data.frame(fav_colors, counts) #make df with above values

#let's plot
ggplot(cols, aes(x = fav_colors, y = counts)) +
  geom_col() +
  coord_flip()
```
 Pretty messy. Luckly, forcats gives us an easy way to reorder our factors so we can better visualize potential patterns/relations between catagories. 

```{r}

cols = cols %>% 
  mutate(fav_colors = fct_reorder(fav_colors, counts, max))

ggplot(cols, aes(x = fav_colors, y = counts)) +
  geom_col() +
  coord_flip()
```


purrr package has lots of useful tools, but one is the ability to nest data and perform functions on each of the nested groups seperately. 

```{r}
library(purrr)

iris_nested = iris %>% 
  group_by(Species) %>% 
  nest()

iris_nested

models = iris_nested %>% 
  mutate(model = lapply(data, function(df) {summary(lm(Petal.Length ~ Petal.Width, data = df))}),
         r_2 = model$r.squared)

models
```
Modelr lets you seamlessly add predictions and residuals as additional columns to an existing data frame
```{r}

library(modelr)

fit = lm(Petal.Length ~ Petal.Width + Sepal.Length + Species, data=iris)

iris = iris %>% 
  add_predictions(fit)

head(iris)
```
We've worked with data frames frequently in the course, but data.table is a powerful package that provides greater flexibility in creating, subsetting, and manipulatiing your data. 
```{r}
library(data.table)

#lets make the df iris a data.table
iris_dt = data.table(iris) 

#now we can use the data.table syntax to select a subset of the species column and compute the mean of the petal length column for just that subset of the species. 

iris_dt[Species == "setosa", .(Petal.Length = mean(Petal.Length))]
```
#patchwork lets you add ggplot plots together in a sensible layout
```{r}
library(patchwork)

p1 = ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_point()

p2 = ggplot(iris, aes(x = Species, y = Petal.Length, fill=Species)) +
  stat_summary(geom="bar") +
  stat_summary(geom="errorbar", width = 0.2) +
  guides(fill=FALSE)

p3 = ggplot(iris, aes(x = Species, y = Petal.Width, fill=Species)) +
  stat_summary(geom="bar") +
  stat_summary(geom="errorbar", width = 0.2) +
  guides(fill=FALSE)

p1 / (p2 + p3)
```


Other useful packages:

shiny - generating simple web apps (https://colinquirk.shinyapps.io/shinyeyes/)

bookdown - creating ebooks with rmarkdown files (https://bookdown.org/cquirk/LetsExploreStatistics/)

blogdown - creating a personal website with r markdown (http://colinquirk.com/)
